The Forward March of Categorical Tolerance in the United States\thanks{Reserved for acknowledgments}

### 2025-06-08 manuscript.tex
Updates from Overleaf

\begin{table}
    \caption{Coefficient Estimates from Zero-Inflated Poisson Models Predicting Categorical Tolerance and Symbolic Exclusion, Joint Prolific and Lucid Samples.}
    \centering
    \includegraphics[trim={0 15cm 45cm 0},clip, width=1.0\textwidth]{Tabs/zinf-reg.png}
    \label{tab:zinf}
\end{table}

Table~\ref{tab:zinf} shows the coefficient estimates from a set of four zero-inflated Poisson regression models. The specifications in each model, for both the count and zero-inflated components, are informed by the descriptive patterns we considered in Figures~\ref{fig:cat-tol} and~\ref{fig:grd-int}. Coefficient estimates corresponding to the count part of the model---predicting the number of dislikes among those who are at risk of expressing them---are shown in the top half of the table. Coefficient estimates for the zero-inflation part of the model, predicting the odds of landing in the categorical tolerant group---and thus not being at risk of engaging in symbolic exclusion---are presented in the lower half of the table. In all models, age, education, income, and political ideology (with higher values indicating a more liberal stance) are entered as ordered numerical inputs. Race (White) and Gender (Female) are entered into the model as dummy indicators. The zero-inflated specifications also contain squared terms for education and income to model expected curvilinear effects (see Figures~\ref{fig:cat-tol-edu} and~\ref{fig:cat-tol-inc}).

Let us begin with the predictors of symbolic exclusion. Model one includes all of the main socio-demographic predictors. The results align with expectations derived from the descriptive patterns considered earlier. Age increases the probability of engaging in symbolic exclusion, but education and income decrease this propensity. Liberals are less likely to engage in symbolic exclusion than conservatives, but white respondents are more likely to express dislikes, compared to non-white respondents. Finally, there is no appreciable gender difference in the propensity to engage in symbolic exclusion via cultural dislikes ($p = 0.52$).

Model 1 includes age as the sole predictor in the zero-inflated part of the model, providing evidence that categorical tolerance is primarily a cultural habit among the young and is less likely to be found among older respondents ($p < 0.01$). Model 2 includes race and gender as additional predictors of categorical tolerance, verifying the expectation that white respondents are less likely to be found in this group (compared to non-white respondents) and women are less likely to be categorical tolerants (compared to respondents who identify as men). Model 3 introduces education as an additional predictor of categorical tolerance, including a linear input for the respondent's education and a square term for the same covariate. The results verify the idea, suggested by Figure~\ref{fig:cat-tol-edu}, the education effect on the odds of being a categorical tolerant is parabolic, with a negative linear component (indicating that lower education person are more likely to be in the categorical tolerant group) and a positive squared component (indicating that at some point the negative effect reverses, and very high-education individuals end up over-represented among the categorical tolerants).

\begin{figure}[ht!]
    \captionsetup[subfigure]{font=footnotesize,labelfont=footnotesize}
    \centering
     \begin{subfigure}[b]{0.3\textwidth}
        \includegraphics[width=1.0\textwidth]{Plots/uni-dist-grd-int-pol.png}
            \caption{Political Ideology}
            \label{fig:grd-int-pol}
    \end{subfigure}
     \begin{subfigure}[b]{0.3\textwidth}
        \includegraphics[width=1.0\textwidth]{Plots/uni-dist-grd-int-edu.png}
            \caption{Education}
            \label{fig:grd-int-edu}
    \end{subfigure}
    \caption{Average number of cultural dislikes across different socio-demographic characteristics.}
    \label{fig:grd-int}
\end{figure}

>>>>

\begin{table}
    \caption{Coefficient Estimates from Zero-Inflated Poisson Models Predicting Categorical Tolerance and Symbolic Exclusion, Joint Prolific and Lucid Samples.}
    \centering
    \includegraphics[trim={0 0cm 45cm 0},clip, width=0.9\textwidth]{Tabs/zinf-reg.png}
    \label{tab:zinf}
\end{table}

Table~\ref{tab:zinf} shows the coefficient estimates from a set of four zero-inflated Poisson regression models. The specifications in each model, for both the count and zero-inflated components, are informed by the descriptive patterns we considered in Figures~\ref{fig:cat-tol} and~\ref{fig:grd-int}. Coefficient estimates corresponding to the count part of the model---predicting the number of dislikes among those who are at risk of expressing them---are shown in the top half of the table. Coefficient estimates for the zero-inflation part of the model, predicting the odds of landing in the categorical tolerant group---and thus not being at risk of engaging in symbolic exclusion---are presented in the lower half of the table. In all models, age, education, income, and political ideology (with higher values indicating a more liberal stance) are entered as ordered numerical inputs. Race (White) and Gender (Female) are entered into the model as dummy indicators. The zero-inflated specifications also contain squared terms for education and income to model expected curvilinear effects (see Figures~\ref{fig:cat-tol-edu} and~\ref{fig:cat-tol-inc}). Finally, I include indicator variables for each of the surveys (one for the 2023 Prolific sample and another for the 2025 Lucid sample), with the older, smaller Prolific survey conducted in 2021 as the reference category.

Let us begin with the predictors of symbolic exclusion. Model one includes all of the main socio-demographic predictors. The results align with expectations derived from the descriptive patterns considered earlier. Age increases the probability of engaging in symbolic exclusion, but education and income decrease this propensity. The education effect, as suggested by Figure~\ref{fig:cat-tol-edu}, is substantively small and not statistically significant ($p = 0.09$) using the traditional cutoff value of $p < 0.05$, and becomes smaller and crosses the threshold to decidedly not statistically significant in later specifications. Politically liberal, on the other hand, has a substantively significant effect, one that is robust, in terms of statistical significance, across models. As suggested by Figure~\ref{fig:cat-tol-pol}, self-identified Liberals are less likely to engage in symbolic exclusion than conservatives. When it comes to ethnoracial identity, self-identified white respondents are more likely to express dislikes compared to respondents who identified with one of the non-white ethnoracial categories ($p < 0.01$). Finally, there is no appreciable gender difference in the propensity to engage in symbolic exclusion via cultural dislikes ($p = 0.52$).

Model 1 includes age as the sole predictor in the zero-inflated part of the model, providing evidence that categorical tolerance is primarily a cultural habit among the young and is less prevalent among older respondents ($p < 0.01$). Model 2 includes race and gender as additional predictors of categorical tolerance, verifying the expectation that white respondents are less likely to be found in this group (compared to non-white respondents) and women are less likely to be categorical tolerants (compared to respondents who identify as men). Model 3 introduces education as an additional predictor of categorical tolerance, including a linear input for the respondent's education and a square term for the same covariate. The results verify the idea, suggested by Figure~\ref{fig:cat-tol-edu}, the education effect on the odds of being a categorical tolerant is parabolic, with a negative linear component (indicating that lower education person are more likely to be in the categorical tolerant group) and a positive squared component (indicating that at some point the negative effect reverses, and very high-education individuals end up over-represented among the categorical tolerants).

That said, the linear component of the education is of only marginal statistical significance in Model 3 ($p = 0.07$), suggesting that the bulk of the systematic effect of education is contained in the positive quadratic component, indicating a larger propensity of the very highly educated to be over-represented among categorical tolerants. This impression is confirmed in Model 4, which adjusts for the non-linear effect of self-reported family income on the odds of being a categorical tolerant. Once we do that, the main negative linear effect of education disappears ($p = 0.22$), but the exponential positive effect remains ($p < 0.05$). Both components of the income effect, on the other hand, have a statistically significant effect, with a negative linear effect and a positive quadratic effect, suggesting an underrepresentation of people in middle-income categories among the categorically tolerant.

\begin{figure}[ht!]
    \captionsetup[subfigure]{font=footnotesize,labelfont=footnotesize}
    \centering
     \begin{subfigure}[b]{0.49\textwidth}
        \includegraphics[width=1.0\textwidth]{Plots/educ-pred-plot.png}
            \caption{Education}
            \label{fig:grd-int-pol}
    \end{subfigure}
     \begin{subfigure}[b]{0.49\textwidth}
        \includegraphics[width=1.0\textwidth]{Plots/inc-pred-plot.png}
            \caption{Income}
            \label{fig:grd-int-edu}
    \end{subfigure}
    \caption{}
    \label{fig:grd-int}
\end{figure}
